---
format: gfm
---

# Simulated Measles Outbreaks in US Cities hosting the 2026 FIFA World Cup

::: {.callout-caution title="Work in progress"}
This scenario simulation is a work in progress. We are still reviewing the model, assumptions, and paramters. Please be mindful of this when interpreting the results.
:::

The following figure summarizes the simulated measles outbreak size over time for the 11 US cities hosting the 2026 FIFA World Cup. Each line represents a single simulation, with the color indicating the vaccination rate in that city. The simulations are based on a mixing model using census data, and the case reporting is as of the latest available date.

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false
library(data.table)
library(ggplot2)

fns <- list.files("scenarios/", pattern = "outbreak.*\\.csv", full.names = TRUE)

# Reading the datasets in
outbreaks <- lapply(fns, \(fn) {
  city <- gsub("scenarios/+([^_]+)_outbreak.+\\.csv", "\\1", fn)
  fread(fn)[, city := city][, virus := NULL][, virus_id := NULL]
}) |> rbindlist()

# Computing the seed
outbreaks[, initial := as.integer(outbreak_size[1]), by = .(city, sim_num)]

# Extracting vaccination rate
vacc_rates <- fread("data/mmr.csv", comment.char = "#")[, .(city, vacc_rate)]

outbreaks <- merge(outbreaks, vacc_rates, by = "city", all.x = TRUE)

# Cases
cases <- fread("data/measles_cases.csv")[, max(date)]
```


```{r}
#| label: plot
#| echo: false
#| warning: false
#| message: false
ggplot(outbreaks, aes(
  x = date, y = outbreak_size,
  group = interaction(city, sim_num)
  )
  ) +
  geom_line(
    aes(color = vacc_rate)
    # color = adjustcolor("steelblue", alpha.f = 0.25)
  ) +
  labs(
    title = "Simulated Measles Outbreak size over time",
    x = "Day",
    y = "Outbreak size",
    # color = "Initial\ncases",
    color = "Vaccination\nrate",
    subtitle = paste(
      "Mixing model with census data.",
      sprintf("Case reporting as of %s", cases)
    ),
    caption = paste(
      "Each line represents a single simulation (60 days).",
      sprintf("Data from %i simulations", length(unique(outbreaks$sim_num)))
    )
  ) +
  facet_wrap(~ city) +
  theme_minimal()
```

You can go over individual city simulations in the `scenarios/` directory:

```{r}
#| label: list-scenarios
#| echo: false
#| results: asis
fns <- sprintf("scenarios/%s.md", unique(outbreaks$city)) |>
  sort()

sprintf("- [%s](%s)", gsub("scenarios/([^_]+).md", "\\1", fns), fns) |>
  cat(sep = "\n")
```