---
format: gfm
params:
  city: "Miami"
  max_pop: 100000
  nthreads: 2
  echo: false
  seed: 8812
  ndays: 60
  nsims: 200
---


# Measles outbreak simulation in `r params$city`

::: {.callout-caution title="Work in progress"}
This scenario simulation is a work in progress. We are still reviewing the model, assumptions, and paramters. Please be mindful of this when interpreting the results.
:::

## Preparing the environment

```{r}
#| label: loading-libraries
#| warning: false
#| message: false
#| echo: false
library(measles)
library(data.table)

if (!exists("params")) {
  params <- list(
    city = "Miami",
    max_pop = 100000,
    nthreads = 4,
    echo = FALSE,
    seed = 8812,
    ndays = 60,
    nsims = 200
    )
}

knitr::opts_chunk$set(
  echo = params$echo,
  warning = FALSE,
  message = FALSE
)

```

Pulling informatou about the city and population from the data files. The mixing matrix is based on the POLYMOD survey data, which provides age-based contact patterns for various cities. We will use the mixing matrix corresponding to the specified city in `params$city`. If the city is not found in the available mixing matrices, an error will be raised.

```{r}
#| label: loading-data
# Population matrix
census_age <- fread("data/census_age.csv")[city == params$city]
population <- fread("data/population.csv")[city == params$city]

# Mixing matrix (age-based, from POLYMOD survey data)
mixing_matrices <- readRDS("data/mixing_matrix.rds")
available_cities <- names(mixing_matrices)

if (!(params$city %in% available_cities)) {
  stop(
    sprintf(
      "Unknown city '%s' in params$city. Available cities in data/mixing_matrix.rds: %s",
      params$city,
      paste(available_cities, collapse = ", ")
    )
  )
}

mixing <- mixing_matrices[[params$city]]

# Retrieving vaccination rate
vacc_rate <- fread(
  "data/mmr.csv",
  comment.char = "#"
  )[city == params$city]$vacc_rate / 100
```

Regarding the number of cases, we leverage information from Johns Hopkins University (JHU) [U.S. Measles Data Repository](https://github.com/CSSEGISandData/measles_data). From it, we pull all the reported cases from the corresponding city. Since the data does not include active cases but reported cases, we will estimate what is the expected number of active cases considering the last reported case as the most recent. Thus, all simulations will have at least one active case at the start, and, if more cases have been reported, they will be included in the simulation.

```{r}
#| label: simulating-active-cases
reported_cases <- fread("data/measles_cases.csv")[city == params$city]
reported_cases <- reported_cases[, .(date)]

# Truncated Geometric distribution
cdf_trunc_geo10 <- \(x) {
  
  p10 <- pgeom(10, prob = 1/4)
  ifelse(x > 10 | x < 0, 1, {
    pgeom(x, prob = 1/4) / p10
  })
  
}

#' Simulating active cases
#' @param dates Vector of dates of class `IDateTime()`
#' @param today Date of class `IDateTime()`, defaulting to the current date.
#' @param cdf CDF function for the truncated Geometric distribution,
#' defaulting to `cdf_trunc_geo10`
#' @return Simulated number of active cases
ractive_cases <- function(
  dates,
  today = max(dates),
  cdf = cdf_trunc_geo10,
  min_cases = 1L
  ) {

  # Day difference
  days_since <- as.integer(today - dates)

  # Probability of being active cases
  p <- 1 - cdf(days_since)

  # Simulating active cases
  max(min_cases, sum(runif(length(p)) < p))

}

# Example (only if interactive)
if (interactive()) {
set.seed(133)
replicate(1000, ractive_cases(reported_cases$date)) |>
  table()
}

# With this in hand, we can then create a function (later) for
# distributing the initial cases using this random sampling approach.
```

```{r}
#| label: checking-size
#| echo: false
#| results: "asis"
if (params$max_pop < sum(population$Population)) {
  cat(
    "::: {.callout-warning}",
    sprintf("The total population in the census data (%i)", sum(population$Population)),
    sprintf("exceeds params$max_pop (%i).", params$max_pop),
    "The population will be scaled down to fit within the specified maximum population size.",
    ":::", sep = "\n"
  )
}
```

```{r}
#| label: adjusting-population
N <- min(params$max_pop, sum(population$Population))

census_age[, total := ceiling(total/sum(total) * N)]
census_age[, vacc_rate := vacc_rate]

N <- sum(census_age$total)
```

```{r}
#| label: calibrating-mixing
calibrate_cm <- function(R0, t_infect) {

  zeros <- rep(0, length(census_age$total))

  obj <- \(s) {
    r_eff <- multigroup.vaccine::vaxrepnum(
      meaninf = 4,
      popsize = census_age$total,
      trmat = mixing * s * t_infect,
      initR = zeros,
      initV = zeros,
      vaxeff = 0
    )

    (r_eff - R0)^2
  }

  stats::optimize(
    f = obj,
    interval = c(0, 1000)
  )$minimum

}

calib_scale <- calibrate_cm(R0 = 12, t_infect = 0.2)
```

## Baseline scenario

```{r}
#| label: creating-model

# Finding hospitalization rate for
# a 10% hospitalization probability
# P(hosp) = hosp_r / (hosp_r + rec_r)
#   => hosp_r = P(hosp) (hosp_r + rec_r)
#             = P(hosp) * hosp_r + P(hosp) * rec_r
#   => hosp_r = P(hosp) * rec_r / (1 - P(hosp))
h_rate <- 0.1 * (1 / 3) / (1 - 0.1)

measles_model <- ModelMeaslesMixing(
  n                            = N,
  prevalence                   = 1,
  transmission_rate            = 0.2,
  vax_efficacy                 = 0.97,
  contact_matrix               = mixing * calib_scale,
  hospitalization_rate         = h_rate,
  hospitalization_period       = 7,
  days_undetected              = 2,
  quarantine_period            = 21,
  quarantine_willingness       = 0.9,
  isolation_willingness        = 0.9,
  isolation_period             = 4,
  prop_vaccinated              = 0.95,
  contact_tracing_success_rate = 0.8,
  contact_tracing_days_window  = 4,
  rash_period                  = 3
)
```


```{r}
#| label: adding-entities
# Adding the entities to the model
add_entities_from_dataframe(
  model = measles_model,
  entities = census_age,
  col_name = "age_group",
  col_number = "total",
  as_proportion = FALSE
)
```

For `r params$city`, the vaccination rate is approximately `r sprintf("%.2f", census_age[, sum(total * vacc_rate)/sum(total)] * 100)`%.

```{r}
#| label: setting-distribution
# Creating the distribution function
dist_fun <- distribute_tool_to_entities(
  prevalence = census_age$vacc_rate,
  as_proportion = TRUE
)

# Setting the distribution function
measles_model |>
  get_tool(0) |>
  set_distribution_tool(dist_fun)
```

For the initial cases, we will use the expected number of active cases based on the reported cases and the sampling approach described above. We will distribute these initial cases randomly across the population.

```{r}
#| label: distribute-initial-cases
# First, we simulate the number of active cases
# outside of the simulation framework.
set.seed(3123)
expected_active_cases <- replicate(
    n = 1000,
    expr = ractive_cases(reported_cases$date)
  ) |>
  mean() |>
  ceiling() |>
  as.integer()

get_virus(measles_model, 0L) |>
  set_distribution_virus(
    distfun = distribute_virus_randomly(
      prevalence = expected_active_cases,
      as_proportion = FALSE
      )
  )
```

We will use `r expected_active_cases` initial cases in the simulations, distributed randomly across the population. We can now run the simulations.

```{r}
#| label: running-simulations
measles_model |>
  run_multiple(
    ndays = params$ndays,
    nsims = params$nsims,
    seed  = params$seed,
    saver = make_saver("outbreak_size", "hospitalizations"),
    nthreads = params$nthreads
  )
```

```{r}
#| label: checking-simulations
summary(measles_model)
```

```{r}
#| label: extracting-results
# Extracting the results
ans <- measles_model |>
  run_multiple_get_results(
    freader = data.table::fread
  )

# Taking a look at the structure
if (params$echo) {
  str(ans)
}

# Converting into data.table format for convenience
outbreak_size <- ans$outbreak_size |> as.data.table()
hospitalizations <- ans$hospitalizations |> as.data.table()
```


```{r}
#| label: plotting-outbreak-size1
# Aggregating to get the final
ggplot2::ggplot(
  outbreak_size,
  ggplot2::aes(x = date, y = outbreak_size)
) +
  ggplot2::geom_point(
    color = "blue",
    alpha = 0.2,
    size = 1.5
  ) +
  ggplot2::labs(
    title = paste("Measles outbreak size in", params$city),
    subtitle = sprintf(
      "Mixing model with census data (%i simulations)",
      params$nsims
    ),
    x = "Day",
    y = "Cases"
  ) +
  ggplot2::theme_minimal()
```


```{r}
#| label: plotting-outbreak-size2
# Aggregating to get the final
final_outbreak <- outbreak_size[date == max(date)]
median_outbreak_size <- median(final_outbreak$outbreak_size)

ggplot2::ggplot(final_outbreak, ggplot2::aes(x = outbreak_size)) +
  ggplot2::geom_histogram(fill = "steelblue", color = "white") +
  ggplot2::labs(
    title = sprintf(
      "Measles outbreak size in %s (median: %.2f)",
      params$city,
      median_outbreak_size
    ),
    subtitle = sprintf(
      "Mixing model with census data (%i simulations)",
      params$nsims
    ),
    x = "Outbreak size",
    y = "Count"
  ) +
  ggplot2::theme_minimal()
```

We can save the results for further analysis.

```{r}
#| label: saving-results
fwrite(outbreak_size, sprintf("%s_outbreak_size.csv", params$city))
fwrite(hospitalizations, sprintf("%s_hospitalizations.csv", params$city))
```

## Session info

This document was generated using the `{measles}` R package version `r packageVersion("measles")` and the `{epiworldR}` package version `r packageVersion("epiworldR")` on `r Sys.Date()`. We used R version `r getRversion()`, running on `r Sys.info()[['sysname']]` with `r Sys.info()[['machine']]` architecture.

