---
format: gfm
---

# Outbreak size vs seed cases with demographic controls

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false
library(data.table)
library(ggplot2)
```

```{r}
#| label: read-simulations
#| echo: false
read_outbreak_files <- function(files, run_label, suffix_pattern) {
  if (length(files) == 0L) {
    return(data.table())
  }

  rbindlist(lapply(files, function(fn) {
    city <- basename(fn)
    city <- sub(suffix_pattern, "", city)

    dat <- fread(fn)[, .(sim_num, date, outbreak_size)]
    dat[, city := city]
    dat[, run_label := run_label]
    dat[, seed_cases := outbreak_size[1], by = .(sim_num)]
    dat
  }))
}

baseline_files <- list.files(
  "scenarios",
  pattern = "_outbreak_size\\.csv$",
  full.names = TRUE
)

plus1_files <- list.files(
  "scenarios",
  pattern = "_outbreak_size_plus1seed\\.csv$",
  full.names = TRUE
)

baseline <- read_outbreak_files(
  baseline_files,
  run_label = "Baseline",
  suffix_pattern = "_outbreak_size\\.csv$"
)

plus1 <- read_outbreak_files(
  plus1_files,
  run_label = "Baseline +1 seed",
  suffix_pattern = "_outbreak_size_plus1seed\\.csv$"
)

outbreaks <- rbindlist(list(baseline, plus1), fill = TRUE)

if (nrow(outbreaks) == 0L) {
  stop("No outbreak size files found in scenarios/.")
}

final_outbreaks <- outbreaks[
  ,
  .SD[date == max(date)],
  by = .(city, sim_num, run_label, seed_cases)
]
```

```{r}
#| label: build-controls
#| echo: false
population <- fread("data/population.csv")[, .(city, Population)]
mmr <- fread("data/mmr.csv", comment.char = "#")[, .(city, vacc_rate)]

census <- fread("data/census_age.csv")
census[, pop_share := total / sum(total), by = city]

age_wide <- dcast(census, city ~ age_group, value.var = "pop_share", fill = 0)
age_matrix <- as.matrix(age_wide[, -"city"])
rownames(age_matrix) <- age_wide$city

age_pca <- prcomp(age_matrix, center = TRUE, scale. = TRUE)
age_controls <- data.table(
  city = rownames(age_matrix),
  age_pc1 = age_pca$x[, 1],
  age_entropy = -rowSums(age_matrix * log(pmax(age_matrix, 1e-12)))
)

analysis <- merge(final_outbreaks, population, by = "city", all.x = TRUE)
analysis <- merge(analysis, mmr, by = "city", all.x = TRUE)
analysis <- merge(analysis, age_controls, by = "city", all.x = TRUE)
analysis <- na.omit(analysis)
```

```{r}
#| label: regression
#| echo: false
fit <- glm(
  outbreak_size ~
    seed_cases +
    log(Population) +
    vacc_rate + age_pc1 + age_entropy,
  family = poisson(),
  data = analysis
)

summary(fit)
```

Regarding the accuracy, the model's pseudo-R² (McFadden's R²) is `r round(1 - fit$deviance / fit$null.deviance, 3)`, indicating that the model explains a modest proportion of the variance in outbreak size. The coefficients suggest that each additional seed case is associated with an increase in expected outbreak size, controlling for population, vaccination rate, and age distribution.

```{r}
#| label: marginal-effect
#| echo: false
avg_city <- data.table(
  seed_cases = mean(analysis$seed_cases),
  Population = mean(analysis$Population),
  vacc_rate = mean(analysis$vacc_rate),
  age_pc1 = mean(analysis$age_pc1),
  age_entropy = mean(analysis$age_entropy)
)

pred_base <- predict(fit, newdata = avg_city, type = "response")
pred_plus1 <- predict(
  fit,
  newdata = transform(avg_city, seed_cases = seed_cases + 1),
  type = "response"
)

marginal_effect <- pred_plus1 - pred_base
marginal_effect
```

Thus, for the average city, the marginal effect of adding one additional case is an increase of approximately `r round(marginal_effect, 2)` in the expected outbreak size.

```{r}
#| label: plot
#| echo: false
ggplot(
  analysis,
  aes(x = seed_cases, y = outbreak_size, color = vacc_rate, size = Population)
) +
  geom_jitter(width = 0.1, alpha = 0.35) +
  facet_wrap(~ run_label) +
  labs(
    title = "Outbreak size versus number of seed cases",
    subtitle = "Point color indicates vaccination rate; point size indicates population",
    x = "Seed cases",
    y = "Final outbreak size",
    color = "Vaccination\nrate (%)",
    size = "Population"
  ) +
  theme_minimal()
```

```{r}
#| label: model-note
#| echo: false
#| results: asis
cat(
  sprintf(
    "Model fitted on %i simulation outcomes across %i cities.",
    nrow(analysis),
    uniqueN(analysis$city)
  )
)
```
