---
format: gfm
---

# idcup

[![Render reports](https://github.com/EpiForeSITE/idcup/actions/workflows/render_reports.yml/badge.svg)](https://github.com/EpiForeSITE/idcup/actions/workflows/render_reports.yml) [![Fetch measles cases](https://github.com/EpiForeSITE/idcup/actions/workflows/update_measles_cases.yml/badge.svg)](https://github.com/EpiForeSITE/idcup/actions/workflows/update_measles_cases.yml)

This project provides agent-based measles outbreak simulations across the 11 major U.S. cities hosting the **2026 FIFA World Cup**. With large international crowds expected, understanding the risk of measles transmission is critical for public-health preparedness. The simulations are built on [epiworldR](https://github.com/UofUEpiBio/epiworldR) and the [measles](https://github.com/UofUEpiBio/measles) R package, using census-derived age structure and MMR vaccination coverage.

For technical details on the repository layout, data sources, model overview, and getting-started instructions, see [DETAILS.md](./DETAILS.md).

# Simulated Measles Outbreaks in US Cities hosting the 2026 FIFA World Cup

::: {.callout-caution title="Work in progress"}
This scenario simulation is a work in progress. We are still reviewing the model, assumptions, and parameters. Please be mindful of this when interpreting the results.
:::

The following figure summarizes the simulated measles outbreak size over time for the 11 US cities hosting the 2026 FIFA World Cup. **The current simulations do not include information about the effect of the World Cup, but merely represent baseline scenarios**. Each line represents a single simulation, with the color indicating the vaccination rate in that city. The simulations are based on a mixing model using census data, and the case reporting is as of the latest available date.

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false
library(data.table)
library(ggplot2)

fns <- list.files("scenarios/", pattern = "outbreak.*\\.csv", full.names = TRUE)

# Reading the datasets in
outbreaks <- lapply(fns, \(fn) {
  city <- gsub("scenarios/+([^_]+)_outbreak.+\\.csv", "\\1", fn)
  fread(fn)[, city := city][, virus := NULL][, virus_id := NULL]
}) |> rbindlist()

# Computing the seed
outbreaks[, initial := as.integer(outbreak_size[1]), by = .(city, sim_num)]

# Extracting vaccination rate
vacc_rates <- fread("data/mmr.csv", comment.char = "#")[, .(city, vacc_rate)]

outbreaks <- merge(outbreaks, vacc_rates, by = "city", all.x = TRUE)

# Cases
cases <- fread("data/measles_cases.csv")[, max(date)]

# Population size
pop <- fread("data/population.csv")
outbreaks <- merge(outbreaks, pop, by = "city")
```


```{r}
#| label: plot
#| echo: false
#| warning: false
#| message: false
ggplot(outbreaks, aes(
  x = date, y = outbreak_size,
  group = interaction(city, sim_num)
  )
  ) +
  geom_line(
    aes(color = vacc_rate),
     alpha = 0.25
    # color = adjustcolor("steelblue", alpha.f = 0.25)
  ) +
  labs(
    title = "Simulated Measles Outbreak size over time (baseline)",
    x = "Day",
    y = "Outbreak size",
    # color = "Initial\ncases",
    color = "Vaccination\nrate",
    subtitle = paste(
      "Mixing model with census data.",
      sprintf("Case reporting as of %s", cases)
    ),
    caption = paste(
      "Each line represents a single simulation (60 days).",
      sprintf("Data from %i simulations", length(unique(outbreaks$sim_num)))
    )
  ) +
  facet_wrap(~ city) +
  theme_minimal()
```

Here is an alternative visualization of the probability that the outbreak size exceeds a given cut-off (10, 20, 50 cases) based on the final outbreak size in the simulations. The points are colored by vaccination rate and sized by population size. The numbers on top of the bars indicate the probability in percentage.

```{r}
#| label: probability
#| warning: false
#| message: false
#| echo: false
#| fig-width: 8
#| fig-height: 10
cuts <- c(10, 20, 50)

# Computing the probability by outbreak size
ob_by_cut <- NULL
ob_final <- outbreaks[date == max(date)]
for (cut in cuts) {
  ob_by_cut <- rbind(
    ob_by_cut,
    ob_final[, .(
      prob = mean(outbreak_size >= cut),
      cut = paste(cut, "+ cases")
      ),
      by = .(city, initial, vacc_rate, Population)
      ]
  )
}

# Preparing the data for plotting
ob_by_cut[, initial := ifelse(initial == 1, "1 initial case", paste(initial, "initial cases"))] 

ob_by_cut[, Population := Population / 1e6]

setorder(ob_by_cut, cut, prob)

ggplot(ob_by_cut[cut != 100], aes(
  x = reorder(city, prob), y = prob, color = vacc_rate)) +
  # geom_bar(stat = "identity") +
  geom_point(aes(color = vacc_rate, size = Population)) +
  geom_segment(
    aes(
      y=0, yend = prob, x=reorder(city, prob),
      xend = reorder(city, prob)
    )
  ) +
  labs(
    title = "Probability of outbreak size (baseline)",
    x = "Outbreak size",
    y = "Probability (of exceeding Outbreak size)",
    subtitle = paste0(
      "Mixing model with census data.\n",
      sprintf("Case reporting as of %s", cases)
    ),
    caption = paste0(
      sprintf(
        "Data from %i simulations",
        length(unique(outbreaks$sim_num))
      )
    ),
    color = "Vaccination\nrate",
    size = "Population\n(millions)"
  ) +
  facet_wrap(~ cut, nrow = 3) +
  coord_flip() +
  theme_minimal() +
  theme(
    text = element_text(size = 20)
  ) + 
  # Adding the numbers on top of the bars
  geom_text(
    aes(
      x = reorder(city, prob), y = prob + .1,
      label = sprintf("%.1f%%", prob * 100)
      ),
    size = 4
  )
```

You can go over individual city simulations in the `scenarios/` directory:

```{r}
#| label: list-scenarios
#| echo: false
#| results: asis
fns <- sprintf("scenarios/%s.md", unique(outbreaks$city)) |>
  sort()

sprintf("- [%s](%s)", gsub("scenarios/([^_]+).md", "\\1", fns), fns) |>
  cat(sep = "\n")
```