---
title: "MMR Coverage from other Sources"
format: gfm
---

In this document, we explore existing school-level MMR coverage data to see if we can find more recent or more granular data than the CDC's state-level estimates. We use data from InsightNet tools, particularly, epiENGAGE's school-level MMR coverage data, which is available for download from their GitHub repository. The single file is downloaded from ForeSITE's GitHub repository, which contains the same information but in a single file.

```{r}
#| label: downloading-vax-data
mmr_schools <- file.path("data", "mmr_schools.csv")

if (!file.exists(mmr_schools)) {
  download.file(
    url = "https://github.com/UofUEpiBio/epiworldRShiny/raw/6a4205c35d647764da2266f6857551de1c291a41/inst/extdata/schools_measles.csv",
    destfile = mmr_schools
  )
}

library(data.table)
mmr_schools <- fread(mmr_schools)
mmr_schools <- mmr_schools[, .(state, county, vaccination_rate)]
mmr_schools <- mmr_schools[, county:= tolower(county)]
```


```{r}
#| label: id-relevant-states
city_lookup <- data.table(
  city = c(
    "Los Angeles", "San Francisco", "New York City", "Boston",
    "Houston", "Dallas", "Philadelphia", "Atlanta",
    "Seattle", "Miami", "Kansas City"
  ),
  state = c(
    "CA", "CA", "NY", "MA",
    "TX", "TX", "PA", "GA",
    "WA", "FL", "MO"
  ),
  state_full = c(
    "California", "California", "New York", "Massachusetts",
    "Texas", "Texas", "Pennsylvania", "Georgia",
    "Washington", "Florida", "Missouri"
  ),
  county = c(
    "Los Angeles", "San Francisco", "New York", "Suffolk",
    "Harris", "Dallas", "Philadelphia", "Fulton",
    "King", "Miami-Dade", "Jackson"
  ) |> tolower()
)
```

Let's now verify which states are covered by this dataset, and which are not. For those that are not, we are accessing new data.

```{r}

# States that have the MMR data
city_lookup[
  which(
    (city_lookup$county %in% unique(mmr_schools$county)) &
      (city_lookup$state %in% unique(mmr_schools$state))
    )
]

# States that do not have the MMR data
city_lookup[
  which(!(
    (city_lookup$county %in% unique(mmr_schools$county)) &
      (city_lookup$state %in% unique(mmr_schools$state))
  ))
]
```

We now compute the average MMR coverage for the states that have data:

```{r}
mmr_schools <- mmr_schools[, .(avg_mmr = mean(vaccination_rate)), by = .(state, county)]

mmr_new <- merge(
  city_lookup,
  mmr_schools,
  by.x = c("state", "county"),
  by.y = c("state", "county"),
  all.y = FALSE
)
```

Comparing with CDC

```{r}
#| label: mmr_from_cdc
mmr_cdc <- fread("data/mmr.csv")

mmr_cdc[, .(city, vacc_rate)] |>
  knitr::kable(digits = 2)

mmr_new[, .(city, avg_mmr = avg_mmr*100)] |>
  knitr::kable(digits = 2)
```


## Immunization in Atlanta, Georgia

The website Immunization reports ([link](https://immunizationstudyreports.s3.us-east-1.amazonaws.com/Child+Report/Child_Report_2023Q1.html#data-by-county)) indicates that the MMR coverage for children aged 19-35 months in Fulton county Georgia is 89.2% for the January-March 2023 quarter. This is higher than the 88.4% reported by the CDC for Atlanta, Georgia.


```{r}
mmr_new <- rbind(
  mmr_new,
  data.table(
    city = "Atlanta",
    state = "GA",
    state_full = "Georgia",
    county = "Fulton",
    avg_mmr = 0.892
  )
)
```

## Immunization in Miami Dade, Florida

According to the Florida Department of Health, the vaccination coverage (full schedule) for 7th grade students in Miami-Dade County for the 2026 was 90.2%, and 90.8% for kindergarteners ([link](https://www.flhealthcharts.gov/ChartsDashboards/rdPage.aspx?rdReport=NonVitalIndNoGrp.Dataviewer)). This is higher than the 88.8% reported by the CDC for Miami, Florida.

```{r}
#| label: mmr_miami
mmr_new <- rbind(
  mmr_new,
  data.table(
    city = "Miami",
    state = "FL",
    state_full = "Florida",
    county = "Miami-Dade",
    avg_mmr = 0.902
  )
)
```


## Immunization in Kansas City, Missouri

In the case of the Jackson county in Missouri, the Missouri Department of Health and Senior Services reports an MMR coverage of 89.5% for Kinder, and 96.2% for 8th grade students in the 2024-2025 school year ([link](https://health.mo.gov/living/families/schoolhealth/dashboard.php)). This is higher than the 88.8% reported by the CDC for Kansas City, Missouri.

```{r}
#| label: mmr_kansas_city
mmr_new <- rbind(
  mmr_new,
  data.table(
    city = "Kansas City",
    state = "MO",
    state_full = "Missouri",
    county = "Jackson",
    avg_mmr = 0.895
  )
)
```

# Final comparison

```{r}
mmr_new[, .(state, county, mmr_new = avg_mmr)]
mmr_cdc[, .(city, vacc_rate)]

# Ensuring caps at the beginning of each word
# for the county name
mmr_new[, county := tools::toTitleCase(county)]

fwrite(mmr_new, file = "data/mmr_extra.csv")
```

